### 2018-11-21 Revision/Paper/method.tex
..

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain \rebut{$\log(A_i)$ values as discussed earlier}.
For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

In Figure~\ref{fig:comp} (left),
we provide a comparison of the square of individual components of the dominant eigenvector, obtained using both, perturbation-based and regression-based estimates of the gradient of the ignition 
delay. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad} \rebut{in both cases}. 

@@

\caption{Left: An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using the two strategies. Right: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
is plotted for successive iterations to illustrate convergence behavior for the regression-based strategy.}

@@

\rebut{Convergence was established once $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ was found
to be smaller than a set tolerance, $\tau$ = 0.05. As shown in Figure~\ref{fig:comp}, 8
iterations were required in both cases. To compare convergence characteristics or the
}

Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
agreement with each other. Moreover, as expected,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is observed 
to decrease with iterations, and requires model evaluations at only 30 samples to converge using $\tau$~=~0.1.

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

We further estimate
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
active subspace is 1-dimensional)
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the perturbation-based and regression-based approaches are plotted in
Figure~\ref{fig:comp_as}.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
%  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the active subspace.}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores based on the two approaches for gradient estimation agree favorably 
with each other as well as those
based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons indicate that the gradient of the ignition delay with respect to the
uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
approaches in this case. Since the number of model evaluations required in the case of
regression are much fewer, this approach seems favorable for the 19-dimensional problem. 
In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.

>>>>

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain \rebut{$\log(A_i)$ values as discussed earlier}.
For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\label{eq:accu}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
to converge in both cases. 
At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
Figure~\ref{fig:comp} (right). Once again, $N$
corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
the same amount of computational effort. However, the perturbation approach is observed to be slightly more
accurate.}

@@

\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
$N$~=~1000 samples)
in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
of the two strategies. Note that the above plots compare the convergence and accuracy of the two
strategies for the same amount of computational effort, quantified by the number of model runs, $M$.
The legend provides the number of samples, $N$ and correspondingly, the
model runs, $M$ required for obtaining a converged dominant eigenvector using the perturbation and
regression strategies. Corresponding value of $N$ and $M$ for the test case are also provided.}}

@@

\rebut{In Figure~\ref{fig:comp} (left), we compare the
squared values of individual components of the dominant eigenvector obtained using the two approaches
at the end of 8 iterations in each case. Additionally, the dominant eigenvector components obtained
for a test case involving $N$~=~1000 samples using the perturbation approach are also plotted. 
The set of eigenvector components for the three cases are found to be in excellent agreement with
each other, indicating that both
approaches have sufficiently converged and are reasonably accurate for this setup.}

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
\caption{\rebut{An illustrative comparison of the SSPs generated using the 
perturbation and the regression strategies for computing the active subspace.}}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}

We further estimate
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 \rebut{since a
1-dimensional active subspace seems reasonably accurate})
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad} (steps 21 and 22).
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the perturbation and regression strategies are plotted in Figure~\ref{fig:comp_as}.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as19D}
%  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{\rebut{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the dominant eigenspace.}}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores based on the two approaches for gradient estimation agree favorably 
with each other as well as those
based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ \rebut{while $A_1$, $A_{15}$, and
$A_{17}$ are also observed to contribute significantly towards its variance.}

The above comparisons indicate that the gradient of the ignition delay with respect to the
uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
approaches in this case. Since the regression approach requires fewer model evaluations for
gradient estimation, it might be a preferred approach for this setting.
In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application \rebut{wherein the activation
energies in the rate law as well as initial pressure, temperature, and stoichiometric conditions
are also considered to be uncertain.}
